Network Configuration Synthesis with Abstract Topologies

### 2016-11-15 paper/abstraction/paper.tex
edits

@@

Second, \sysname can scale substantially better than its nearest competitor Propane that operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, \sysname can be two orders of magnitude
faster than Propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

In this work, we focus on synthesizing configurations for a network that runs the Border Gateway Protocol (BGP), the standard for routing across independent organizations and inside large data centers. It scales effectively and allows operators to define flexible policies.

We provide a brief background on BGP focused on aspects relevant to our work. How a BGP-based network carries traffic depends on how the routers exchange and process routing announcements. To invite traffic from a neighbor to a destination, routers announce the destination's address {\em prefix} (e.g., 10.1.1.0/24\footnote{An IP address
such as 10.1.1.0 is 32 bits, with each of the four values representing 8 bits. The prefix 10.1.1.0/24 denotes the set of addresses that share the first 24 bits with 10.1.1.0, \IE, all addresses that begin with 10.1.1.}) to the neighbor. Traffic flows in the opposite direction to such route announcements.

@@

When a BGP router receives one or more announcements for a prefix from different neighbors, it selects which one to use (if any) based on its configuration. This selection can depend on the AS path, communities, or the neighbor that sent the announcement, and it can even discard certain announcements (i.e., never use them for sending traffic). Then, based again on its configuration, the router forwards the selected route to one or more neighbors, after adding itself to the AS path, and optionally, modifying communities.

@@

When announcing destinations to certain neighbors, a BGP router may perform {\em route aggregation}, \IE, announce a single prefix that covers multiple prefixes (e.g., announce 10.1.0.0/16 instead of 10.1.[1--255].0/24). The covering prefix is announced as long as the router has a valid route for {\em any} of the covered prefixes. Aggregation helps reduce the memory consumption of routers, but it can lead to traffic black holes~\cite{route-aggregation}. A router that announces 10.1.0.0/16, because it has a route to 10.1.1.0/24, may also get traffic for 10.1.2.0/24 to which it has no route.

@@

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ say that, within any pod P, the number of outgoing edges from a node in the $TG$ role to a node in the $AG$ role equals the number of nodes in the $AG$ role. Similarly, the number of outgoing edges from a node in the $AG$ role to a node in the $TG$ role equals the number of nodes in the $TG$ role. These constraints capture the fact that, within any pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure that, within pods P and Q, the $AG$ and $TG$ roles have the same number of routers, which is less than the number of routers in the spine role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example, by growing the network from Figure~\ref{fig:example} to Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies, potentially at the expense of analysis precision. We also include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$~simply to show that constraints do not have to be in the form of inequalities. Operators can use logical formulas from any theory supported by modern SMT solvers.

@@

% The \text{} is needed for some reason for formatting
\para{Synthesis} \sysname generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on Propane (for concrete networks); we show that it can be extended to abstract inputs and that the extensions are correct.

@@

%\ryan{Possibly cut here}
%As an example, suppose we want to compute the number of of disjoint paths between any global ToR and spine router in Figure~\ref{fig:example3}.
%The analysis begins with the fact
%$S_P S_{TG} (1,1)$---there is some group of
%nodes of size 1 in the $TG$ role in some pod $P$ such that any initial
%node in the $TG$ role has 1 disjoint path to it (and all such paths are
%edge-disjoint). Since the abstract topology specifies that each node in
%the $TG$ role is in
%a full mesh with nodes in the $AG$ role within any given pod, we
%infer the fact $S_P A_{AG}(e1,1)$. This says that
%any group of nodes of size e1 in the $AG$ role is reachable with at
%least 1 path to each such node from any initial node in
%the $TG$ role.
%
%After one more step, the analysis will learn a fact of the form $A_{S}(1,1)$, which means that any single spine node is reachable via a single disjoint path from any initial global ToR node
%In fact, due the the valley-free routing constraint, a single path is
%the worst-case fault-tolerance level for the concrete network from
%Figure~\ref{fig:example2}.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, $P$) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $P \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $P(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $P(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\tilde{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

@@

\begin{defn}
\label{def:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete product graph $PG = (G',\mathit{start},P)$ and abstract product graph $PG^A = (G'^A, \mathit{start}^A, P^A)$, then we can construct a new homomorphism $f_{pg} : G' \rightarrow G'^A$ in the following way:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{defn}
%
As we will see in \S\ref{sec:generation}, we have proven that our generation strategy commutes with template instantiation, meaning that we obtain the same results if we instantiate the abstraction early, or if we defer the instantiation until after template generation.
%
Definition~\ref{def:homomorphism} provides intuition as to why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.

@@

The possibility of network failures exacerbates the difficulty of constructing correct configurations.
Link failures in networks occur frequently; it is not uncommon for a large network to experience dozens of failures in any given day~\cite{dc-failure-study}. However, existing tools~\cite{condor,propane} reason about fault-tolerance only for concrete topologies. The problem for \sysname is harder because it must verify that all possible concrete instantiations of an abstract topology satisfy the fault-tolerance policy, which requires a fundamentally different approach.

%
We frame satisfying the fault-tolerance policy provided by the user as an analysis problem over the structure of the PG. In particular, we develop an analysis that uses information about the structure of  concrete instantiations of the abstract topology to infer bounds on the number of edge-disjoint paths between pairs of concrete nodes.

%
For each node in the abstract PG,
the idea is to maintain invariants learned about the number of disjoint paths
to groups of concrete routers in the node's role.
More specifically, we maintain invariants of the form:
%
\[ \begin{array}{c}
  L_{X_1}, \ldots, L_{X_n}(j,k)
\end{array} \]
\noindent
%
where each label $L \in \{S,A\}$ is either $S$, which stands for ``some" or is $A$, which stands for ``all". There is one label for each pod in the abstraction pod hierarchy under which the abstract node appears. For a given node, $L_{X_1}$ corresponds to the outermost pod, $L_{X_{n-1}}$ corresponds to the innermost pod, and $L_{X_n}$ to the the node itself.
Semantically, $L_{X_1}, \ldots, L_{X_n}(j,k)$ means that starting from some concrete source node, for some/all pods $X_1, \ldots, X_{n-1}$
and for some/all groups of nodes in the role $X_n$ of size $j$,
%, $L_{X_2}, \ldots, L_{X_n}(j,k)$ holds. The base case $L_{X_n}(j,k)$ means that for "some"/``all" groups of size $j$ concrete routers,
there are $k$ paths to each such that all $j*k$ paths are edge-disjoint.

For example,
an inference of the form $A_{Q} A_{TL}(2,3)$ states that, from some fixed source location, for \emph{all} pods Q, and \emph{all} groups of 2 nodes in the TL role, there are 6 disjoint paths to the group---3 for each of the 2 nodes.

@@

To learn about disjoint paths, we develop a collection of sound inference rules for invariants of the form above. Figure~\ref{fig:inference-rules} shows each rule. They are read from bottom to top. The label on the bottom left is a known invariant. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to invariants learned after applying the inference rule. The adjacent box shows the conditions that must be \emph{valid}, given the abstraction constraints, for the rule to apply. Validity ensures that the condition holds for all concrete instantiations of the topology.

In general, the inference rules try to learn about both the largest number of disjoint paths to any single node in an abstract role as well as the largest reachable group in a particular role with at least one disjoint path to each.

The first rule, \infrule{I-out1} applies to a learned invariant of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we cannot exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$.
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

@@

\subsection{An inference algorithm}

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any invariant learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Invariants learned with $j=0$ or $k=0$ are discarded.

@@

Recall the policy for global prefixes in the data center.
%
\begin{code}
\$GP \Path \End(TG) \ensuremath{\cap}
       \Novalley(\{TG,TL\},\{AG,AL\},\{S\}) \ensuremath{\cap}
       !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows part of the abstract PG representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies each inference rule to this initial fact. Rules \infrule{I-mesh1} together with \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles results in a call to $\nu$Z to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The algorithm will then eventually apply \infrule{I-out2} to learn that any single spine node is reachable at $(S,0)$. It will also apply \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at $(S,0)$ and that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The next step is to use \infrule{I-mesh2} together with \infrule{I-local} to infer than any single node in the TL role for any pod Q is reachable via at least 2 disjoint paths. This process will continue until a fixed point is reached.

@@

The first idea is to tag every BGP announcement with a community that encodes the state of each DFA when advertising a route to a neighbor. Routers then allow announcements that correspond to an edge in the PG and block others.
For example, in the concrete PG in figure~\ref{fig:example-compilation}, router S1 will allow an advertisement from its neighbors N1 and N2. If it uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ before sending this to its neighbors. If it uses neighbor N2, then it will add the tag $(0,1)$ instead.

@@

\begin{defn}
Let $m \geq_{rank} n$ be a relation over PG vertices that holds iff $P(n) = \emptyset$ or $\min P(m) \geq \min P(n)$. Intuitively $m \geq_{rank} n$ means that paths ending at PG node $m$ are no better than paths ending at $n$.
%
\end{defn}
\noindent
%

%\begin{defn}
The PG can be viewed as a labelled transition system by pushing the location from each directed edge's target node onto the edge. That is, $m\overset{l}{\rightarrow}n$ if edge $(m,n) \in E'$ and $\tilde{n} = l$.
%\end{defn}
%
For example, in the concrete PG we would have the transition $(S1,1,0)\overset{A2}{\rightarrow}(A2,1,0)$.

\begin{defn}
We write $m \leq m'$ if the subgraph reachable from $m$ and $m'$ respectively form a simulation relation with respect to the $\geq_{rank}$ relation.
\end{defn}

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ in state $(1,0)$ must be preferred to those received at $S1$ in state $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds.

@@

Suppose an operator wants to expand the concrete data center by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

However, template variables tell us exactly where this kind of change can happen. The only way in which the spine routers depend on anything non-local is in the match on prefix template variables. Therefore, we can prevent modifications to the spine routers' configurations by associating a new unique community tag for template variable {\small \$GP} and then replacing each match on {\small \$GP} with a test on the new community tag. The new ToR must now simply attach the new community tag to advertisements for its prefixes so downstream routers can match against it. Figure~\ref{fig:bgp-configs} (right) shows the templates after the transformation.

@@

\sysname is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.

Although the disjoint path analysis takes place over the PG, each application of the inference rules from figure~\ref{fig:inference-rules} depends only on the topology. Therefore, we lazily apply the rules and cache the satisfiability and minimization calls to $\nu$Z after their first use. Furthermore, the cached results are shared across different prefixes, each of which may have a unique PG representation.

@@

We use a fat tree topology and scale the size of the topology according the the number of pods $k$. The abstract topology uses one abstract node for each tier of the data center with an additional node for separate local and global ToRs.

>>>>

\newcommand{\Topo}[1]{\ensuremath{{\mathrm{topo}}(#1)}}
\newcommand{\Pref}{\ensuremath{\mathrm{pref}}}

\newcommand{\Conserv}{\cal{C}}

@@

Second, \sysname can scale substantially better than its nearest competitor Propane, which operates over concrete topologies. As the number of devices in a
network grows, the number of distinct roles will often stay constant~\cite{surgery}.
Consequently, \sysname can be two orders of magnitude
faster than Propane, taking less than 10 seconds to synthesize templates for large structured networks.

@@

In this work, we focus on synthesizing configurations for a network that runs the Border Gateway Protocol (BGP), the standard for routing across independent organizations and inside large data centers. It scales effectively and allows operators to define flexible policies. We provide a brief background on BGP focused on aspects relevant to our work.

The way in which a BGP-based network forwards traffic depends on how the routers exchange and process \emph{routing announcements}. To invite traffic from a neighbor to a destination, routers announce the destination's address {\em prefix} (e.g., 10.1.1.0/24\footnote{An IP address
such as 10.1.1.0 is 32 bits, with each of the four values representing 8 bits. The prefix 10.1.1.0/24 denotes the set of addresses that share the first 24 bits with 10.1.1.0, \IE, all addresses that begin with 10.1.1.}) to the neighbor. Traffic flows in the opposite direction to such route announcements.

@@

When a BGP router receives one or more announcements for a prefix from different neighbors, it selects which one to use (if any) based on its configuration. This selection can depend on the AS path, communities, or the neighbor that sent the announcement, and it can even discard certain announcements (\IE, never use them for sending traffic). Then, based again on its configuration, the router forwards the selected route to one or more neighbors, after adding itself to the AS path, and optionally, modifying communities.

@@

When announcing destinations to certain neighbors, a BGP router may perform {\em route aggregation}, \IE, announce a single prefix that covers multiple prefixes (\EG, announce 10.1.0.0/16 instead of 10.1.[1--255].0/24). The covering prefix is announced as long as the router has a valid route for {\em any} of the covered prefixes. Aggregation helps reduce the memory consumption of routers, but it can lead to traffic black holes~\cite{route-aggregation}. A router that announces 10.1.0.0/16, because it has a route to 10.1.1.0/24, may also get traffic for 10.1.2.0/24 to which it has no route.

@@

For example, in Figure~\ref{fig:example3}, the first two constraints: $e1 = AG$ and $e2 = TG$ say that, within any pod P, the number of outgoing edges from a node in the $TG$ role to a node in the $AG$ role equals the number of nodes in the $AG$ role. Similarly, the number of outgoing edges from a node in the $AG$ role to a node in the $TG$ role equals the number of nodes in the $TG$ role. These constraints capture the fact that, within any pod, the global aggregators and ToRs are in a full mesh. Furthermore, the constraints $AG = AL$ and $AG \leq S$ ensure that, within pods P and Q, the $AG$ and $TG$ roles have the same number of routers, which is less than or equal to the number of routers in the spine role $S$.
%
The constraint $e3 \geq 2$ says that, in each pod, each aggregator node has at least 2 outgoing edges to nodes in the spine role. Symmetrically, the constraint $e4 \geq 1$ says that, for each pod P, each node in the spine role has at least one outgoing edge to a node in the $AG$ role. Similar constraints appear for the local aggregator role. The constraint $2 \leq S \leq 4$ makes explicit the possibility for growth, for example, by growing the network from Figure~\ref{fig:example} to Figure~\ref{fig:example2}. In general, we need not bound the number of spine routers to admit more concrete topologies, potentially at the expense of analysis precision. We also include the constraints $(S \mod AG) = 0$, and $(S \mod AL) = 0$~simply to show that constraints do not have to be in the form of inequalities. Operators can use logical formulas from any theory supported by modern SMT solvers.

@@

% The \text{} is needed for some reason for formatting
\para{Synthesis} \sysname generates templates from the inputs above in three phases. First, it combines the abstract topology and routing policy into a {\em product graph} (PG) that compactly captures the flow of routing information over the topology in a policy-compliant manner. The algorithm for this phase builds on Propane (for concrete networks); we show that it can be extended to abstract inputs.
%and that the extensions are correct.

@@

%\ryan{Possibly cut here}
%As an example, suppose we want to compute the number of of disjoint paths between any global ToR and spine router in Figure~\ref{fig:example3}.
%The analysis begins with the fact
%$S_P S_{TG} (1,1)$---there is some group of
%nodes of size 1 in the $TG$ role in some pod $P$ such that any initial
%node in the $TG$ role has 1 disjoint path to it (and all such paths are
%edge-disjoint). Since the abstract topology specifies that each node in
%the $TG$ role is in
%a full mesh with nodes in the $AG$ role within any given pod, we
%infer the fact $S_P A_{AG}(e1,1)$. This says that
%any group of nodes of size e1 in the $AG$ role is reachable with at
%least 1 path to each such node from any initial node in
%the $TG$ role.
%
%After one more step, the analysis will learn a fact of the form $A_{S}(1,1)$, which means that any single spine node is reachable via a single disjoint path from any initial global ToR node
%In fact, due the valley-free routing constraint, a single path is
%the worst-case fault-tolerance level for the concrete network from
%Figure~\ref{fig:example2}.

@@

For each regular path constraint $r_i$ from $r_1 ~\Prefer~ .. ~\Prefer~ r_k$, we construct a DFA for the reverse of $r_i$. A DFA for $r_i$ is defined as a tuple ($\Sigma, Q_i, F_i, q_{0_i}, \sigma_i$). The alphabet $\Sigma$ is the set of topology locations, $Q_i$ is the set of states for automaton i, $F_i$ is the set of final states, $q_{0_i}$ is the initial state, and $\sigma_i \colon Q_i \times \Sigma \rightarrow Q_i$ is the state transition function.
%
The PG is a tuple ($G'$, $start$, \Pref) where $G' = (V',E')$ has
vertices $V' \colon V \times Q_1 \times \dots \times Q_j$,
edges $E' \colon V' \times V'$,
a unique starting vertex $start$,
and a preference function $\Pref \colon V' \rightarrow 2^{\set{1, \dots, j}}$, mapping nodes in the PG to a set of path ranks.

The PG is constructed by adding an edge from state $m = (l_m, q_{m_1}, \dots, q_{m_k})$ to $n = (l_n, q_{n_1}, \dots, q_{n_k})$ whenever $\sigma_i(q_{m_i}, l_n) = q_{n_i}$ for each $i$ and $(l_m,l_n) \in E$.
%
We add edges from the $\mathit{start}$ node to any $m = (l, q_{m_1}, \dots, q_{m_k})$ when $\sigma_i(q_{0_i}, l) = q_{m_i}$ for each $i$.
%
The preference function $\Pref(m)$ denotes the rank of paths through the PG ending at node $m$ and is defined as $\Pref(m) = \set{i~\vert~q_{m_i} \in F_i}$.
%
Finally, we write $\Topo{m} = l$ to extract the topology location from a PG node, when $m = (l, q_{m_1}, \dots, q_{m_k}) \in V'$.

@@

\begin{lem}
\label{lem:homomorphism}
If we have a graph homomorphism $f : G \rightarrow G^A$, concrete product graph $PG = (G',\mathit{start},\Pref)$ and abstract product graph $PG^A = (G'^A, \mathit{start}^A, \Pref^A)$, then there is a homomorphism $f_{pg} : G' \rightarrow G'^A$ where:
\[ \begin{array}{rcl}
  f_{pg}( \mathit{start} ) & = & \mathit{start}^A  \\
  f_{pg}( (l,q_1,\ldots,q_n) ) & = & (f(l),q_1,\ldots,q_n) \\
\end{array} \]
\end{lem}
%
As shown in \S\ref{sec:generation}, our generation strategy commutes with template instantiation, meaning that we obtain the same results if we instantiate the abstraction early, or if we defer the instantiation until after template generation.
%
%Lemma~\ref{lem:homomorphism} provides intuition as to why our strategy for generating templates is correct.
%Abstract product graphs are related to concrete product graphs via a graph homomorphism, just as abstract topologies are related to concrete topologies via a graph homomorphism.

@@

The possibility of network failures exacerbates the difficulty of constructing correct configurations.
Link failures in networks occur frequently; it is not uncommon for a large network to experience dozens of failures in any given day~\cite{dc-failure-study}. However, existing tools~\cite{condor,propane} reason about fault-tolerance only for concrete topologies. \sysname provides substantially stronger guarantees: all possible concrete instantiations of an abstract topology satisfy the fault-tolerance policy.

%The problem for \sysname is harder because it must verify that all possible concrete instantiations of an abstract topology satisfy the fault-tolerance policy, which requires a fundamentally different approach.

%
We frame satisfying the fault-tolerance policy provided by the user as an analysis problem over the structure of the PG. In particular, we develop an analysis that uses information embedded in the abstract topology to infer bounds on the number of edge-disjoint paths between pairs of concrete nodes.

%
For each node in the abstract PG,
the idea is to infer facts learned about the number of disjoint paths
to groups of concrete routers in the node's role.
More specifically, we maintain fact of the form:
%
\[ \begin{array}{c}
  L_{X_1}, \ldots, L_{X_n}(j,k)
\end{array} \]
\noindent
%
where each label $L \in \{S,A\}$ is either $S$, which stands for ``some" or is $A$, which stands for ``all". There is one label for each pod in the abstraction pod hierarchy under which the abstract node appears. For a given node, $L_{X_1}$ corresponds to the outermost pod, $L_{X_{n-1}}$ corresponds to the innermost pod, and $L_{X_n}$ to the node itself.
Semantically, $L_{X_1}, \ldots, L_{X_n}(j,k)$ means that starting from some concrete source node, for some/all pods $X_1, \ldots, X_{n-1}$
and for some/all groups of nodes in the role $X_n$ of size $j$,
%, $L_{X_2}, \ldots, L_{X_n}(j,k)$ holds. The base case $L_{X_n}(j,k)$ means that for "some"/``all" groups of size $j$ concrete routers,
there are $k$ paths to each such that all $j*k$ paths are edge-disjoint.

For example,
an inference of the form $A_{Q} A_{TL}(2,3)$ states that, from the given source location, for \emph{all} pods Q, and \emph{all} groups of 2 nodes in the TL role, there are 6 disjoint paths to the group---3 for each of the 2 nodes.

@@

Figure~\ref{fig:inference-rules} displays the collection of rules used to infer facts about disjoint paths.
Each rule is read from bottom to top. The label on the bottom left is a known fact. We use $L$ to represent a rule that is parametric over the label ($S$ or $A$). The labels on other nodes correspond to facts learned after applying the inference rule. The adjacent box shows the conditions that must be valid, given the abstraction constraints, for the rule to apply. 
%Validity ensures that the condition holds for all concrete instantiations of the topology.

Some of the inference rules (\EG, \infrule{I-out2} and \infrule{I-mesh2}) try to learn about the largest number of disjoint paths to any single node in an abstract role, while others (\EG, \infrule{I-out1} and \infrule{I-mesh1}) try to learn about the largest reachable group in a particular role with at least one disjoint path to each node in that group. Both kinds of rules are useful.

The first rule, \infrule{I-out1} applies to a learned fact of the form $L_m(j,k)$ where the number of outgoing edges from any concrete node in the $m$ role is greater than 0. In the worst case, the largest group of concrete nodes we could hope to reach at the $n$ role would be $e_1$ since all j nodes at the bottom may have outgoing edges to the same concrete nodes at the top. Furthermore, the total number of disjoint paths to the $j$ nodes at the bottom is equal to $j*k$. Since extending the existing disjoint paths with disjoint edges keeps the paths disjoint, and since we cannot exceed the current number of disjoint paths to the concrete nodes in role $m$ on the bottom, the largest reachable group for the role $n$ on the top will be $min(j*k,e_1)$.
%
We conservatively use $1$ for the number of disjoint paths to each node in role $n$, since when $n$ is very large, all reachable nodes in role $m$ might only have a single edge to completely different nodes in $n$.
%However, because we do not know for certain that all $j$ nodes in role $m$ have outgoing edges to the same nodes at the top, we conservatively use $1$ for the number of disjoint paths to each node on top.
%For example, if $n$ is very large, then each node in role $m$ may have only a single edge to each reachable node in role $n$.

@@

\subsection{Inference algorithm}

The abstract disjoint path analysis starts from a fixed source location $src$ and repeatedly tries to apply every inference rule from Figure~\ref{fig:inference-rules} until it reaches a fixed point. The algorithm applies an inference rule when the rule's condition is valid given the abstract topology constraints. Because the inference rules may continue to yield larger and larger symbolic expressions, we make the following observation to ensure termination: for any invariant learned of the form $L(j,k)$, it is sound to instead infer $L(j',k')$ if $j' \leq j$ and $k' \leq k$. Therefore, for each inference $L(j,k)$ we minimize the symbolic expressions for $j$ and $k$ subject to the topology constraints using the optimizing SMT solver $\nu$Z~\cite{z3opt}. Facts learned with $j=0$ or $k=0$ are discarded.

@@

Recall the policy for global prefixes in the data center.
%
\begin{code}
\$GP \Path \End(TG) \ensuremath{\cap}
       \Novalley(\{TG,TL\},\{AG,AL\},\{S\}) \ensuremath{\cap}
       !(\Enter(Peer) & \Exit(Peer))
\end{code}
\noindent
%
Figure~\ref{fig:example-inference} shows part of the abstract PG representation for this routing policy. The inference algorithm starts from the node $(TG,0,0)$ with the initial fact $S_P S_{TG}(1,\infty)$ (\IE, no restriction on the number of disjoint paths initially).
%
The first step applies each inference rule to this initial fact. The algorithm uses rules \infrule{I-mesh1} and \infrule{I-local} to reason about connectivity within a single pod for the TG and AG roles. It makes a call to $\nu$Z to minimize the expression $min(\infty, AG)$, which results in 2. Therefore, the algorithm learns a new invariant of the form $S_P A_{AG} (2,1)$ for node $(AG, 0)$ to indicate that in some pod P, any group of 2 nodes is reachable.
%
The algorithm will then eventually apply \infrule{I-out2} to learn that any single spine node is reachable at $(S,0)$. It will also apply \infrule{I-striping} to determine that there is some group of at least 2 spine routers reachable at $(S,0)$ and that there is some group of at least 2 nodes reachable in the AL role in state $(AL,0)$.

@@

The next step is to use \infrule{I-mesh2} together with \infrule{I-local} to infer that any single node in the TL role for any pod Q is reachable via at least 2 disjoint paths. This process will continue until a fixed point is reached.

@@

The first idea is to tag every BGP announcement with a community that encodes the state of each DFA when advertising a route to a neighbor. Routers then allow announcements that correspond to an edge in the PG and block others.
For example, in the concrete PG in Figure~\ref{fig:example-compilation}, router S1 will allow an advertisement from its neighbors N1 and N2. If it uses the path advertised by neighbor N1, then it will add the tag $(1,0)$ before sending this to its neighbors. If it uses neighbor N2, then it will add the tag $(0,1)$ instead.

@@

\begin{defn}
Let $m \geq_{rank} n$ be a relation over PG vertices that holds iff $\min (\Pref(m)) \geq \min (\Pref(n))$ or $\Pref(n) = \emptyset$. Intuitively $m \geq_{rank} n$ means that paths ending at node $m$ are no better than paths ending at $n$.
%
\end{defn}
\noindent
%

%\begin{defn}
The PG can be viewed as a labelled transition system by pushing the location from each directed edge's target node onto the edge. That is, $m\overset{l}{\rightarrow}n$ if edge $(m,n) \in E'$ and $\Topo{n} = l$.
%\end{defn}
%
For example, in the concrete PG we would have the transition $(S1,1,0)\overset{A2}{\rightarrow}(A2,1,0)$.

\begin{defn}
We write $m \leq m'$ if the subgraph reachable from $m$ and $m'$ respectively form a simulation relation with respect to $\geq_{rank}$.
\end{defn}

If we determine that each router has a total order with respect to the simulation relation ($\leq$) for its corresponding PG nodes, then we can simply sort the nodes and prefer advertisements from peers of node $m$ over those of $m'$ whenever $m' \leq m$. This condition ensures that BGP devices will converge to the most preferred paths in the network even when failures occur.
%
For example, in the concrete PG the inequality $(S1,0,1) \leq (S1,1,0)$ holds since nodes on the left side of the PG can always match transitions made on the right hand side with respect to the $\geq_{rank}$ relation. This relation does not hold the other way around since $(S1,0,1) \ngeq_{rank} (S1,1,0)$. Therefore, advertisements received at $S1$ with tag $(1,0)$ must be preferred to those received at $S1$ with tag $(0,1)$.
%
Notice that in the abstract PG $(S,0,1) \leq (S,1,0)$ also holds.

@@

Suppose an operator wants to expand the concrete data center from Figure~\ref{fig:example} by adding an additional ToR router to the TG role. Per the network routing policy, the new ToR router will advertise any owned prefixes provided by looking up {\small \$GP} in $\Gamma$. Because the new topology matches the abstraction, the compiled templates will remain the same. However, in the spine router configurations, the match on the global prefix template variable {\small \$GP} must be expanded when concretizing the template to include the new prefixes added by the ToR. Unfortunately, this small change to the topology results in a change to every single spine router configuration.

However, template variables tell us exactly where this kind of change can happen. The only way in which the spine routers depend on anything non-local is in the match on prefix template variables. Therefore, we can prevent modifications to the spine routers' configurations by associating a new unique community tag for template variable {\small \$GP} and then replacing each match on {\small \$GP} with a test on the new community tag. The new ToR must now simply attach the new community tag to advertisements for its prefixes so downstream routers can match against it. Figure~\ref{fig:bgp-configs} (right) shows several templates after the transformation.

@@

\sysname is implemented in approximately 9500 lines of F\# code. The compiler accepts both concrete and abstract topologies and will generate configurations for open-source Quagga~\cite{quagga} routers by default. The fault-tolerance analysis uses $\nu$Z~\cite{z3opt} to both test validity and minimize variables subject to the topology constraints. Since the analysis typically calls the SMT solver many times with relatively small optimization problems, we use a timeout of 200ms.

Although the disjoint path analysis takes place over the PG, each application of the inference rules from Figure~\ref{fig:inference-rules} depends only on the topology. Therefore, we lazily apply the rules and cache the satisfiability and minimization calls to $\nu$Z after their first use. Furthermore, the cached results are shared across different prefixes, each of which may have a unique PG representation.

@@

We use a fat tree topology and scale the size of the topology according the number of pods $k$. The abstract topology uses one abstract node for each tier of the data center with an additional node for separate local and global ToRs.
